Phase Transitions in Landscape Connectivity

### 2020-08-29 writing/manuscript.tex
critical pig

\[P(V_j|V_i)=\frac{f(d_{ij}, \alpha)}{\sum_k f(d_{ik},\alpha)}\]

@@

\[dN_i=\lambda_i N_i \big(1 - \frac{N_i}{K_i}\big)dt+\sigma_pN_idW\]

Here, \(N_i\) is the abundance of population \(L_i\), \(\lambda_i\) is the
strength of density dependence in that population, and \(K_i\) is the
carrying capacity of \(L_i\). \(\sigma_p\) represents that standard
deviation in abundance due to local stochasticity as a proportion of
\(N_i\). Here \(\sigma_p\) represents an amalgamation of all factors
contributing to local stochasticity in population dynamics, although it
should be noted that the relative contribution of different factors to
local stochasticity can drive significant variation in dynamics
\cite{melbourne_extinction_2008}. For the sake of reducing the size of parameter
space, here we consider all populations as having the same \(\lambda_i\)
and \(K_i\), however, future work could include exploring the
source-sink dynamics in this system by varying intrinsic growth rates
and carrying capacities across populations.

We use an SDE representation because they have been used to study phase
transitions in stochastic systems before \cite{brock_early_2012, kuehn_mathematical_2011}, and they have many nice properties for inference. SDEs
have been used to study extinction dynamics before, as it is relatively
straightforward to compute the mean time until extinction (MTE) using
the Kolmogorov Backward Equation (@lande\_stochastic\_2003).

\hypertarget{diffusion-on-spatial-graphs}{%
\subsubsection{Diffusion on Spatial
Graphs}\label{diffusion-on-spatial-graphs}}

When we model a landscape with a spatial graph, we have to decide how
different nodes affect one another. The processes that connect
landscapes are inherently stochastic. The probability that an individual
migrates within its lifetime, \(p_m\), and where it goes, \(P(V_j|V_i)\)
, are both stochastic processes. Under some conditions, we can
effectively model stochastic dynamics across space using diffusion
models. In essence, a diffusion model assumes that at each timestep the
system will change according to the expected value of stochastic
dispersal. Diffusion models have seen widespread use in ecology and
other fields (@ovaskainen\_empirical\_2008,
@holmes\_partial-differential\_1994, @okubo\_diffusion\_2011).

Given our dispersal potential \(P(V_j|V_i)\), we can define the
diffusion matrix \(\Phi\) as

\[\Phi_{ij} = \begin{cases} 1 - p_m \quad\quad &\text{if}\ i = j \\ P(V_j|V_i)p_m & \text{if}\ i \neq j\end{cases}\]

where \(\Phi_{ij}\) represents the probability that any individual born
in \(V_i\) reproduces in \(V_j\).

Now we move to considering the dynamics of the abundance of the
population at \(V_i\), denoted \(N_i\). We can now represent the
dynamics due to diffusion of this system as

\[\dot{N_i}=(1-p_m)N_i+ \sum_j p_mP(V_j|V_i)N_j\]

@@

\[\frac{d\vec{N}}{dt}=\Phi^T\vec{N}\]

@@

\[\frac{d\vec{N}}{dt} = g(\Phi^T \vec{N})\]

@@

\begin{equation} \label{diffusion_model}
  d \vec{N} = {I} (\Phi^T \vec{N})(\vec{k}-\vec{N}) dt + \sigma_p \vec{k} \ d\vec{W}
\end{equation}

@@

As described above, measures of correlation in space and time are often
used in the study of phase transitions. When the order parameter gains
or loses correlation in space and time is used as an indicator of when a
system changes qualitative phases. Within the context of ecology, the
crosscorrelation function, \(CC\), has long been used as a measure of
synchronous dynamics (@liebhold\_spatial\_2004). Here, with a subdivided
population, we consider the mean crosscorrelation compared across all
populations, \({PCC}=\frac{1}{n^2}\sum_{i,j} CC(V_i,V_j)\) . As an
example, in \ref{async_and_sync}, panel (a) show an example of low PCC,
and panel (b) shows an example of high PCC.

\begin{figure}[h]
\caption{The abundance of 5 populations (each in a different color) across time. Panel (a):  These populations show low synchrony, $PCC=0.1$. Panel (b): These populations show high sychrony, $PCC=0.89$ }

\label{async_and_sync}

@@

\begin{figure}
    \includegraphics[width=15cm]{figs/figure4.pdf}
    \caption{}
    \label{}
\end{figure}

>>>>

\[\Phi_{ij} = P(V_j|V_i)=\frac{f(d_{ij}, \alpha)}{\sum_k f(d_{ik},\alpha)}\]

@@

\[dN_i=\lambda_i N_i \bigg(1 - \frac{N_i}{K_i}\bigg)dt+\sigma N_idW\]

Here, \(N_i\) is the abundance of population \(L_i\), \(\lambda_i\) is the
strength of density dependence in that population, and \(K_i\) is the
carrying capacity of \(L_i\). \(\sigma_p\) represents that standard
deviation in abundance due to local stochasticity as a proportion of
\(N_i\). Here \(\sigma\) represents an amalgamation of all factors
contributing to local stochasticity in population dynamics, although it
should be noted that the relative contribution of different factors to
local stochasticity can drive significant variation in dynamics
\cite{melbourne_extinction_2008}.

We can write this system as the matrix equation
\begin{equation}
        d\boldsymbol{N} = \boldsymbol{I} \boldsymbol{\lambda} \bigg(\boldsymbol{I}(\boldsymbol{K} - \boldsymbol{N})\boldsymbol{K}^{-1} \bigg) d\boldsymbol{t} + \boldsymbol{I} \boldsymbol{\sigma} \boldsymbol{N} d\boldsymbol{W}
\end{equation}

For the sake of reducing the size of parameter
space, here we consider all populations as having the same \(\lambda_i\)
and \(K_i\), however, future work could include exploring the
source-sink dynamics in this system by varying intrinsic growth rates
and carrying capacities across populations.

We use an SDE representation because they have been used to study phase
transitions in stochastic systems before \cite{brock_early_2012, kuehn_mathematical_2011}, and they have many nice properties for inference. SDEs
have also been used in ecology to study extinction dynamics before, as it is relatively straightforward to compute the mean time until extinction (MTE) using
the Kolmogorov Backward Equation \cite{lande_stochastic_2003}.

\hypertarget{diffusion-on-spatial-graphs}{%
\subsection{Diffusion on Spatial
Graphs}\label{diffusion-on-spatial-graphs}}

When we model a landscape with a spatial graph, we have to decide how
different nodes affect one another. The processes that connect
landscapes are inherently stochastic. The probability that an individual
migrates within its lifetime, $m$, and where it goes, $\Phi_{ij}$, are both stochastic . Under some conditions, we can
effectively model stochastic dynamics across space using diffusion
models. In essence, a diffusion model assumes that at each timestep the
system will change according to the expected value of stochastic
dispersal. Diffusion models have seen widespread use in ecology and
other fields \cite{ovaskainen_empirical_2008, holmes_partial-differential_1994, okubo_diffusion_2011}.

Given our dispersal potential $\Phi_{ij}$, we can define the
diffusion matrix \( D \) as

\[D_{ij} = \begin{cases} 1 - m \quad\quad &\text{if}\ i = j \\ \Phi_{ij}m & \text{if}\ i \neq j\end{cases}\]

We can now represent the
dynamics due to diffusion of this system as

\[\dot{N_i}=(1-m)N_i+ \sum_j \Phi_{ij}mN_j\]

@@

\[\frac{d\vec{N}}{dt}=D^T\vec{N}\]

@@

\[\frac{d\vec{N}}{dt} = g(D^T \vec{N})\]

@@

\begin{equation} \label{diffusion_model}
  d \boldsymbol{N} = \boldsymbol{I} (D^T \boldsymbol{N})\bigg(\boldsymbol{I}\boldsymbol{K}^{-1}(\boldsymbol{K}-D^T \boldsymbol{N})\bigg) d\boldsymbol{t} + \boldsymbol{I}\boldsymbol{\sigma} \boldsymbol{N} \ d\boldsymbol{W}
\end{equation}

@@

As described above, measures of correlation in space and time are often
used in the study of phase transitions. When the order parameter gains
or loses correlation in space and time is used as an indicator of when a
system changes qualitative phases. Within the context of ecology, the
crosscorrelation function, \(CC\), has long been used as a measure of
synchronous dynamics \cite{liebhold_spatial_2004}. Here, with a subdivided
population, we consider the mean crosscorrelation compared across all
populations, \({PCC}=\frac{1}{n^2}\sum_{i,j} CC(V_i,V_j)\) . As an
example, in \ref{async_and_sync}, panel (a) show an example of low PCC,
and panel (b) shows an example of high PCC.

\begin{figure}[h]
\includegraphics[width=15cm]{./figs/figure2}
\caption{The abundance of 2 populations (top) and 8 populations (bottom) across time. Get the values of synchrony }
\label{async_and_sync}

@@

\pagebreak
\begin{figure}
    \includegraphics[width=15cm]{figs/figure3}
    \caption{}
    \label{}
\end{figure}

\begin{figure}
    \includegraphics[width=15cm]{figs/figure4}
    \caption{}
    \label{}
\end{figure}

\begin{figure}
    \includegraphics[width=15cm]{figs/figure5}
    \caption{}
    \label{}
\end{figure}
